Modelling BCG vaccination in the UK

### 2019-05-03 index/05-evidence-policy-change.Rmd
add thesis links

The Bacillus Calmette–Guérin (BCG) vaccine remains the only licensed vaccine for use against Tuberculosis (TB), although its use globally is controversial due to evidence of variable effectiveness,[@Mangtani2014a] and waning protection 10-15 years after vaccination.[@Abubakar2013] Global usage of the BCG varies between no vaccination, universal vaccination, and high-risk group vaccination and may target either neonates or school-aged children.[@Pilger2012b; @Zwerling2011a] The World Health Organization (WHO) recommends vaccination for all neonates as early as possible after birth in high burden settings, with vaccination in low burden settings being dependent on the country specific epidemiology of TB.[@WHO2017] This recommendation is based on the strong evidence that the BCG is highly protective in children,[@Rodrigues1993; @Colditz1994] whilst its effectiveness has been shown to vary with latitude when given later in life.[@Mangtani2014]

@@

I considered the original assumption of a 9% annual decrease in incidence rates as well as three scenarios based on those considered by the JVCI BCG subgroup:[@JVCIBCG2002; @JVCIBCG2003] these were a 3.9% decrease, a 1.9% decrease, and no change annual in incidence rates. I also estimated the annual decrease in incidence rates in the ethnic White UK born using two proxy measures. The first proxy measure was the annual change in notifications in England and Wales, which was estimated using data from Public Health England (PHE). The standard deviation (SD) of this measure was then calculated using the `prop.test` function in R.[@R] The second proxy used was the annual decrease in the UK born age-specific incidence rates in the English population. These were calculated using notification data from the Enhanced Tuberculosis surveillance system (ETS) and the June Labour Force Survey.[@PHE2017] Incidence rates (with SDs) were estimated using the epiR package.[@EpiR]. Uncertainty was incorporated by sampling from a normal distribution for both proxy measures. Data collection for the ETS began in 2000, I therefore estimated incidence rates between 1984 and 1999, and for the years between notifications surveys (1974-1977 and 1979-1892), using local regression for each sample fitted to the incidence rates published in [@Sutherland1989], and the estimated incidence rates from 2000 on-wards. For years prior to 1973 the annual decreases were assumed to be the mean of the previous 3 years of data. For both proxy measures the annual decreases in incidence rates post 2016 were assumed to be the average of the estimates in 2013, 2014, and 2015.

@@

This Chapter has reassessed some of the evidence previously used in decision making, updating the approach with new data. However, as 15 years of detailed surveillance data have been collected since the ending of the BCG schools' scheme it is now possible to use regression-based approaches to estimate the direct impact on incidence rates of ending the BCG schools' scheme. These approaches could also be used to estimate the impact of vaccinating high-risk neonates, which may outweigh any negative impacts of ending the BCG schools' scheme. In addition, the development, and use, of a transmission dynamic model would allow the more accurate estimation of indirect effects and the forecasting of long-term impacts.

>>>>

The Bacillus Calmette–Guérin (BCG) vaccine remains the only licensed vaccine for use against Tuberculosis (TB), although its use globally is controversial due to evidence of variable effectiveness,[@Mangtani2014a] and waning protection 10-15 years after vaccination.[@Abubakar2013] Global usage of the BCG varies between no vaccination, universal vaccination, and high-risk group vaccination and may target either neonates or school-aged children.[@Pilger2012b; @Zwerling2011a] The World Health Organization (WHO) recommends vaccination for all neonates as early as possible after birth in high burden settings, with vaccination in low burden settings being dependent on the country specific epidemiology of TB.[@WHO2017] This recommendation is based on the strong evidence that the BCG is highly protective in children,[@Rodrigues1993; @Colditz1994] whilst its effectiveness has been shown to vary with latitude when given later in life (see Chapter \@ref(background)).[@Mangtani2014]

@@

I considered the original assumption of a 9% annual decrease in incidence rates as well as three scenarios based on those considered by the JVCI BCG subgroup:[@JVCIBCG2002; @JVCIBCG2003] these were a 3.9% decrease, a 1.9% decrease, and no change annual in incidence rates. I also estimated the annual decrease in incidence rates in the ethnic White UK born using two proxy measures. The first proxy measure was the annual change in notifications in England and Wales, which was estimated using data from Public Health England (PHE). The standard deviation (SD) of this measure was then calculated using the `prop.test` function in R.[@R] The second proxy used was the annual decrease in the UK born age-specific incidence rates in the English population. These were calculated using notification data from the Enhanced Tuberculosis surveillance system (ETS) and the June Labour Force Survey (see Chapter \@ref(data)).[@PHE2017] Incidence rates (with SDs) were estimated using the epiR package.[@EpiR]. Uncertainty was incorporated by sampling from a normal distribution for both proxy measures. Data collection for the ETS began in 2000, I therefore estimated incidence rates between 1984 and 1999, and for the years between notifications surveys (1974-1977 and 1979-1892), using local regression for each sample fitted to the incidence rates published in [@Sutherland1989], and the estimated incidence rates from 2000 on-wards. For years prior to 1973 the annual decreases were assumed to be the mean of the previous 3 years of data. For both proxy measures the annual decreases in incidence rates post 2016 were assumed to be the average of the estimates in 2013, 2014, and 2015.

@@

This Chapter has reassessed some of the evidence previously used in decision making, updating the approach with new data. However, as 15 years of detailed surveillance data have been collected since the ending of the BCG schools' scheme it is now possible to use regression-based approaches to estimate the direct impact on incidence rates of ending the BCG schools' scheme (see Chapter \@ref(direct-eff)). These approaches could also be used to estimate the impact of vaccinating high-risk neonates, which may outweigh any negative impacts of ending the BCG schools' scheme. In addition, the development (see Chapter \@ref(model-development)), and use, of a transmission dynamic model would allow the more accurate estimation of indirect effects and the forecasting of long-term impacts (see Chapter \@ref(model-results)).
